---
title: "Covid91 vaccine study Final2021"
author: "Brennen Long"
date: "5/3/2021"
output: 
  html_document:
    number_sections: true
    toc: true


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
#separate by gender
male <- subset(FinalData, sex=="guy")
female <- subset(FinalData, sex=="gal")
#separate by drug use
yes <- subset(FinalData, DrugUser=="yes")
no <- subset(FinalData, DrugUser=="no")
#separate by sexual orientation
straight <- subset(FinalData, LGBTQ=="straight")
gay <- subset(FinalData, LGBTQ=="gay")
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Gender

### Males

```{r}
graphMale <- male %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
basicC <- ggplot(graphMale,aes(x=infected,y=count,fill=treatment)) 
basicC + geom_bar(stat="identity",position="dodge") + ggtitle("Male subgroup \n count")
basicCC <- ggplot(graphMale,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge") + ggtitle("Male subgroup \n percent")
```

This research points to the fact that the vaccine can be harmful. 

```{r}
maleTable <- xtabs(~treatment + infected, data=male)
rowPerc(maleTable)
colPerc(maleTable)
```

People with COVID-19 is small in both the placebo and vaccinated groups.  More than half of the people with the COVID-19 vaccine were vaccinated, while more than half of the people recieved the placebo.  This also points that the vaccine is harmful.

```{r}
chisq.test(maleTable)
chisqtestGC(maleTable)
fisher.test(maleTable)
```

The alternate hypothesis is that infection and treatment have some correlation. With a low p- value at 0.0005801, the null hypothesis (that infection and treatment are unrelated) is rejected. In fact, vaccinating males seems to make them more vulnerable to COVID-19.

### Females

```{r}
graphFemale <- female %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
basicC <- ggplot(graphFemale,aes(x=infected,y=count,fill=treatment)) 
basicC + geom_bar(stat="identity",position="dodge") + ggtitle("Female subgroup \n count")
basicCC <- ggplot(graphFemale,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge") + ggtitle("Female subgroup \n percent")
```

The drug seems to be far more effective on females.

```{r}
femaleTable <- xtabs(~treatment + infected, data=female)
rowPerc(femaleTable)
colPerc(femaleTable)
```

```{r}
chisq.test(femaleTable)
chisqtestGC(femaleTable)
fisher.test(femaleTable)
```

The chi-squared statistic here (66.42) is far greater than in the male subgroup (11.720), which would imply a greater effectiveness on females

## Drug Use

### Drug users

```{r}
graphDrug <- yes %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
basicC <- ggplot(graphDrug,aes(x=infected,y=count,fill=treatment)) 
basicC + geom_bar(stat="identity",position="dodge") + ggtitle("Drug user subgroup \n count")
basicCC <- ggplot(graphDrug,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge") + ggtitle("Drug user subgroup \n percent")
```  

The vaccine seems to be very effective on drug users, although this data doesn't differentiate what drugs are being used. This would need further study to analyze.

```{r}
drugTable <- xtabs(~treatment + infected, data=yes)
rowPerc(drugTable)
colPerc(drugTable)
```

In both the "covid" and "OK" categories, those who were given the vaccine, had a greater percentage (96.37) that were okay than who were given the placebo (93.54).  

```{r}
chisq.test(drugTable)
chisqtestGC(drugTable)
fisher.test(drugTable)
```

A p-value between 0.0023 and 0.003 (both less than 0.05) suggests that the null hypothesis is rejected. The odds ratio of 0.5455 is reliable since the upper barrier (0.8194) is less than 1. 

## Sexual Orientation

### Gays

```{r}
graphGay <- gay %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
basicC <- ggplot(graphGay,aes(x=infected,y=count,fill=treatment)) 
basicC + geom_bar(stat="identity",position="dodge") + ggtitle("Gay subgroup \n count")
basicCC <- ggplot(graphGay,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge") + ggtitle("Gay subgroup \n percent")
```

The vaccine is incredibly ineffective here; more than 60% of the gays who got Covid were given the drug. This could be because of a larger influx of gay people getting the vaccine.

```{r}
gayTable <- xtabs(~treatment + infected, data=gay)
rowPerc(gayTable)
colPerc(gayTable)
```

Covid is clearly associated with the vaccine, and the placebo generally results in the patient not contracting it. Thus, this vaccine is harmful to those in the LGBTQ community, for an unknown reason.  

```{r}
chisq.test(gayTable)
chisqtestGC(gayTable)
fisher.test(gayTable)
```

The odds ratio is more than 1 and p- value is much less than 0.05, meaning that the null hypothesis (that infection and treatment are not correlated) can be rejected. An odds ratio of ~2.4922 means that those in the LGBTQ community are much more likely to get Covid than expected.

# Conclusion

The vaccine is harmful to to males by making them more vulnerable to COVID-19 after recieving the vaccine, and same with the LGBT community.  On the other hand, women and drug users alike had a greater effectiveness with the vaccine.  So we can firmly establish that men and the LGBT community are more at risk for getting the vaccine and women and drug users benefit from the vaccine.